\documentclass[11pt]{amsart}
\usepackage[margin=1.2in,marginparsep=0.1in,marginparwidth=1in]{geometry}
\usepackage{amssymb,amsmath,amsthm,amstext,amscd,latexsym,graphics,graphicx,bbm,caption}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[plainpages=false,colorlinks=true, linkcolor=blue, citecolor=red, pagebackref]{hyperref}
\usepackage{enumitem}


\makeatletter
\def\@settitle{\begin{center}%
		\baselineskip14\p@\relax
		\bfseries
		\uppercasenonmath\@title
		\@title
		\ifx\@subtitle\@empty\else
		\\[1ex]\uppercasenonmath\@subtitle
		\footnotesize\mdseries\@subtitle
		\fi
	\end{center}%
}
\def\subtitle#1{\gdef\@subtitle{#1}}
\def\@subtitle{}
\makeatother


%% LaTeX Definitions
%\newcounter{countup}

\newcommand{\rup}[1]{\lceil{#1}\rceil}
\newcommand{\rdown}[1]{\lfloor{#1}\rfloor}
\newcommand{\ilim}{\mathop{\varprojlim}\limits} % inverse limit
\newcommand{\dlim}{\mathop{\varinjlim}\limits}  % direct limit
\newcommand{\surj}{\twoheadrightarrow}
\newcommand{\inj}{\hookrightarrow}
\newcommand{\tensor}{\otimes}
\newcommand{\ext}{\bigwedge}
\newcommand{\Intersection}{\bigcap}
\newcommand{\Union}{\bigcup}
\newcommand{\intersection}{\cap}
\newcommand{\union}{\cup}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% new new commands :) %%%%%%%%%%%%%%%%
\newcommand{\supp}{{\rm Supp}}
\newcommand{\Exceptional}{{\rm Ex}}
\newcommand{\del}{\partial}
\newcommand{\delbar}{\overline{\partial}}
\newcommand{\boldphi}{\mbox{\boldmath $\phi$}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\udiv}{\underline{\Div}}

%%%%%%%%%%%%%%%%%

\newcommand{\Proj}{{\P roj}}
\newcommand{\sEnd}{{\sE nd}}
\newcommand{\mc}{\mathcal}
\newcommand{\mb}{\mathbb}
\newcommand{\an}{{\rm an}} 
\newcommand{\red}{{\rm red}}
\newcommand{\codim}{{\rm codim}}
\newcommand{\Dim}{{\rm dim}}
\newcommand{\rank}{{\rm rank}}
\newcommand{\Ker}{{\rm Ker  }}
\newcommand{\Pic}{{\rm Pic}}
\newcommand{\per}{{\rm per}}
\newcommand{\ind}{{\rm ind}}
\newcommand{\Div}{{\rm Div}}
\newcommand{\Hom}{{\rm Hom}}
\newcommand{\Aut}{{\rm Aut}}
\newcommand{\im}{{\rm im}}
\newcommand{\Spec}{{\rm Spec \,}}
\newcommand{\Sing}{{\rm Sing}}
\newcommand{\sing}{{\rm sing}}
\newcommand{\reg}{{\rm reg}}
\newcommand{\Char}{{\rm char}}
\newcommand{\Tr}{{\rm Tr}}
\newcommand{\Gal}{{\rm Gal}}
\newcommand{\Min}{{\rm Min \ }}
\newcommand{\Max}{{\rm Max \ }}
\newcommand{\Alb}{{\rm Alb}\,}
\newcommand{\Mat}{{\rm Mat}}
%\newcommand{\GL}{{\rm GL}\,}        % For the general linear group
\newcommand{\GL}{{\G\L}}
\newcommand{\Ho}{{\rm Ho}}
\newcommand{\ie}{{\it i.e.\/},\ }
\renewcommand{\iff}{\mbox{ $\Longleftrightarrow$ }}
\renewcommand{\tilde}{\widetilde}
% Skriptbuchstaben
\newcommand{\sA}{{\mathcal A}}
\newcommand{\sB}{{\mathcal B}}
\newcommand{\sC}{{\mathcal C}}
\newcommand{\sD}{{\mathcal D}}
\newcommand{\sE}{{\mathcal E}}
\newcommand{\sF}{{\mathcal F}}
\newcommand{\sG}{{\mathcal G}}
\newcommand{\sH}{{\mathcal H}}
\newcommand{\sI}{{\mathcal I}}
\newcommand{\sJ}{{\mathcal J}}
\newcommand{\sK}{{\mathcal K}}
\newcommand{\sL}{{\mathcal L}}
\newcommand{\sM}{{\mathcal M}}
\newcommand{\sN}{{\mathcal N}}
\newcommand{\sO}{{\mathcal O}}
\newcommand{\sP}{{\mathcal P}}
\newcommand{\sQ}{{\mathcal Q}}
\newcommand{\sR}{{\mathcal R}}
\newcommand{\sS}{{\mathcal S}}
\newcommand{\sT}{{\mathcal T}}
\newcommand{\sU}{{\mathcal U}}
\newcommand{\sV}{{\mathcal V}}
\newcommand{\sW}{{\mathcal W}}
\newcommand{\sX}{{\mathcal X}}
\newcommand{\sY}{{\mathcal Y}}
\newcommand{\sZ}{{\mathcal Z}}
% Sonderbuchstaben mit Doppellinie
\newcommand{\A}{{\mathbb A}}
\newcommand{\B}{{\mathbb B}}
\newcommand{\C}{{\mathbb C}}
\newcommand{\D}{{\mathbb D}}
\newcommand{\E}{{\mathbb E}}
\newcommand{\F}{{\mathbb F}}
\newcommand{\G}{{\mathbb G}}
\newcommand{\HH}{{\mathbb H}}
\newcommand{\I}{{\mathbb I}}
\newcommand{\J}{{\mathbb J}}
\newcommand{\M}{{\mathbb M}}
\newcommand{\N}{{\mathbb N}}
\renewcommand{\O}{{\mathbb O}}
\renewcommand{\P}{{\mathbb P}}
\newcommand{\Q}{{\mathbb Q}}
\newcommand{\R}{{\mathbb R}}
\newcommand{\T}{{\mathbb T}}
\newcommand{\U}{{\mathbb U}}
\newcommand{\V}{{\mathbb V}}
\newcommand{\W}{{\mathbb W}}
\newcommand{\X}{{\mathbb X}}
\newcommand{\Y}{{\mathbb Y}}
\newcommand{\Z}{{\mathbb Z}}
\newcommand{\Sh}{\sS h}
\newcommand{\deltaop}{\Delta^{op}(\sS h(Sm/\mathbf{k}))}
\newcommand{\pdeltaop}{\Delta^{op}(P\sS h(Sm/\mathbf{k}))}
%\newcommand{\psh}{\pi_0^{\text{\tiny pre}}}
\newcommand{\psh}{\pi_0}
\renewcommand{\k}{\mathbf{k}}

\newcommand{\colim}{{\rm colim \,}}
\newcommand{\DM}[2]{\mathbf{DM}_{#2}^{\mathit{eff}}(#1)}

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{question}[theorem]{Question}



\begin{document}
% Document information
\title{Seminar on Moduli Theory}
\subtitle{\today}
\author{Neeraj Deshmukh}
%\date{\today}
%\address[IISERM]{Indian Institute of Science Education and Research, Mohali}

\maketitle

\section{Plan}

One way to go to through these sessions is go recall theory of schemes (at level of things in Hartshorne chapter two), by doing a beeline through all the definitions, properties, etc. However, I feel since the point of doing this exercise is to become more comfortable in working with scheme, we will just do lots of examples instead. By this, I mean we will just to prove some things in very concrete situations. This will help you build a concrete picture of the generalities.

List of some things to discuss (just do lots of examples):

\begin{enumerate}
	\item Definition of a scheme.
	\item Say affine communication lemma \textit{stress this!}
	\item examples
	\begin{enumerate}
		\item $\P^n$ and it sheaf theory! This already clarifies the $\sO(n)$'s
	\end{enumerate}
\end{enumerate}


Things to say: finite-generation, valuative criteria for $\P^1$ (using DVR's - how giving a point and it's specialisation is the same as giving a map $\Spec R\rightarrow \P^1$).\\
List of examples:
\begin{enumerate}
	\item $\A^1$ with a double point (what are quasi-coherent sheaves on this?)
	\item $\A^2\setminus \lbrace 0,0\rbrace$ (what is the structure sheaf?).
	\item $\P^1$ (its structure sheaf!).
	\item $V_{+}(x^2+y^2+z^2)$ over $\R$ and $\C$. The point is that there is a change of coordintes which 
	\item Blow-up of $\A^2$ at a point. (because everyone should know about blow-ups!)
	\item $\Spec R[x_1,x_2,\ldots]$ as an example of something non-noetherian.
	\item An example of a scheme without a closed point.
\end{enumerate}

List of morphisms:
\begin{enumerate}
	\item $x\mapsto x^2$ (more, generally $x^n$). This covers ramified, finitely presented, flat.
	\item a non-quasi-compact open-immersion. Polynomial ring in infinitely many variables and knock off the origin. Also, the origin of is an example of something 
\end{enumerate}

\section{Right at the beginning..}
For the sake of completeness we begin by reviewing the definition of a locally ringed space.

\begin{definition}
	\label{definition-locally-ringed-space}
	Locally ringed spaces.
	\begin{enumerate}
		\item A {\it locally ringed space $(X, \mathcal{O}_X)$}
		is a pair consisting of a
		topological space $X$ and a sheaf of rings $\mathcal{O}_X$ all of whose stalks
		are local rings.
		\item Given a locally ringed space $(X, \mathcal{O}_X)$ we say that
		$\mathcal{O}_{X, x}$ is the {\it local ring of $X$ at $x$}.
		We denote $\mathfrak{m}_{X, x}$ or simply $\mathfrak{m}_x$
		the maximal ideal of $\mathcal{O}_{X, x}$. Moreover, the
		{\it residue field of $X$ at $x$} is the residue field
		$\kappa(x) = \mathcal{O}_{X, x}/\mathfrak{m}_x$.
		\item A {\it morphism of locally ringed spaces}
		$(f, f^\sharp) : (X, \mathcal{O}_X) \to (Y, \mathcal{O}_Y)$
		is a morphism of ringed spaces such that for all $x \in X$
		the induced ring map $\mathcal{O}_{Y, f(x)} \to \mathcal{O}_{X, x}$ is a
		local ring map.
	\end{enumerate}
\end{definition}


We know that affine schemes are locally ringed spaces: we take $\Spec R$ with the zariski topology and for any principal open set $D(f)$ we assign the ring $R_f$. So, any ring $R$ produces the sheaf $\tilde{R}$ on $\Spec R$. This is called the tilde construction. (sanity check: if you can do this, then you should be able to construct a sheaf on $\Spec R$ for any $R$-module $M$).


\begin{definition}
	A {\it scheme} is a locally ringed space with the property that
	every point has an open neighbourhood which is an affine scheme.
	A {\it morphism of schemes} is a morphism of locally
	ringed spaces. The category of schemes will be denoted
	$Sch$.
\end{definition}

\begin{definition}
	Let $(X,\mathcal{O}_X)$ be a scheme. A sheaf of modules on $X$ is a sheaf $\mathcal{F}$ on $X$ such that for every open set $U$, $\mathcal{F}(U)$ is an $\mathcal{O}_X(U)$-module. We say that a sheaf of modules $\mathcal{F}$ is \textit{quasi-coherent} if for every affine open $U\simeq \Spec(R)$, the sheaf $\mathcal{F}|_U$ on $U$ is of the form $\tilde{M}$ for some $R$-module $M$.
\end{definition}


Make special note of the next lemma. This basically lets us reduce problems about schemes to statement about affine schemes (ergo, ring theory), whenever the problem at hand is of a \textit{local} nature. Ravi Vakil calls this \textit{affine communication lemma}.


\begin{lemma}
	\label{lemma-locally-P}
	Let $X$ be a scheme. Let $P$ be a local property of rings.
	The following are equivalent:
	\begin{enumerate}
		\item The scheme $X$ is locally $P$.
		\item For every affine open $U \subset X$ the property
		$P(\mathcal{O}_X(U))$ holds.
		\item There exists an affine open covering $X = \bigcup U_i$ such that
		each $\mathcal{O}_X(U_i)$ satisfies $P$.
		\item There exists an open covering $X = \bigcup X_j$
		such that each open subscheme $X_j$ is locally $P$.
	\end{enumerate}
	Moreover, if $X$ is locally $P$ then every open subscheme
	is locally $P$.
\end{lemma}

This is how commutative algebra meets geometry. Often, the properties that we want to consider are ``globalised" versions of statements about rings. Some examples, normality, reduced, etc.\footnote{You can also ``globalise" morphisms of rings, but now you have two schemes to work locally on. We'll do this soon.}

Examples
\begin{enumerate}
	\item Two ways of gluing $\A^1\setminus\lbrace 0\rbrace$: $x\mapsto x$ or $x\mapsto 1/x$.
	\begin{enumerate}
		\item Double origin: What are global sections? what are quasi-coherent sheaves?
		\item $\P^1$: What are global sections?
	\end{enumerate}
	\item A normal scheme, a reduced scheme, Noetherian scheme.
\end{enumerate}


%\bibliography{Mybib.bib}
\bibliographystyle{alpha}


\end{document}



